Reproduce Analyses Reported in "Phenotypic evolution is more constrained in simpler food webs

### 2019-01-10 analyses/reproduce_analyses.Rmd
moved old evolution simulations on the adaptive landscape to a separate .Rmd file

---
title: "Supplementary Material"
author: "Matthew A. Barbour"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
#output:
#  pdf_document:
#    df_print: paged
#    fig_caption: true
#    citation_package: natbib
#    includes:  
#      in_header: preamble-latex.tex
#bibliography: references_ECD_supp_mat.bib
fontsize: 12pt
---

@@

## Load required libraries ----
library(evolqg) # need to load before dplyr
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(kableExtra)
library(stringr)
library(MVN)        # assessing multivariate normality assumption

@@

Reproduce Figure 5 in main manuscript.

@@

```{r }
# Plot helpers
AF_range <- bind_rows(RF_Clutch.Pref$absolute_fitness, RF_Diam.Clutch$absolute_fitness, RF_Diam.Pref$absolute_fitness) %>%
  summarise(min = min(average), max = max(average))

# Size x Preference
AF_plot_Diam.Pref <- RF_Diam.Pref$absolute_fitness %>% 
  #filter(Foodweb == "Complex") %>%
  ggplot(., aes(x=sc.Diam, y=sc.sqrt.Pref)) +
  geom_raster(aes(fill=average)) +
  #stat_contour(aes(z = average), breaks = breaks["Complex"], linetype = "dotted", color = "black") +
  #geom_raster(data = filter(RF_Diam.Pref$absolute_fitness, Foodweb == "Simple"), aes(fill=average)) +
  #stat_contour(data = filter(RF_Diam.Pref$absolute_fitness, Foodweb == "Simple"), aes(z = average), breaks = breaks["Simple"], linetype = "dotted", color = "black") +
  facet_wrap(~Foodweb) +
  xlab("Chamber diameter (SDs)") +
  ylab("Oviposition preference (SDs)") +
  scale_x_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_fill_viridis(name = "Mean larval\nsurvival", trans="log", breaks = my_breaks, labels = my_breaks, limits = c(AF_range$min, AF_range$max)) +
  theme(strip.background = element_blank(),
        strip.text.x=element_text(margin=margin(b=4)))

# Clutch x Preference
AF_plot_Clutch.Pref <- RF_Clutch.Pref$absolute_fitness %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=sc.log.Clutch)) +
  geom_raster(aes(fill=average)) +
  facet_wrap(~Foodweb) +
  xlab("Oviposition preference (SDs)") +
  ylab("Clutch size (SDs)") +
  scale_x_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_fill_viridis(name = "Mean larval\nsurvival", trans="log", breaks = my_breaks, labels = my_breaks, limits = c(AF_range$min, AF_range$max)) +
  theme(strip.background = element_blank(),
        strip.text.x=element_text(margin=margin(b=4)))

Remove 2nd contour line

```{r Figure-Multivariate-Landscapes-2, echo=FALSE, fig.cap="Fitness landscapes of gall traits in complex vs. simple food webs. Each panel corresponds to a different combination of traits: clutch size and gall diameter (A); clutch size and Oviposition preference (B); Oviposition preference and gall diameter (C). Note that traits for all plots range 1 SD below and above the mean (=0)."}

# Format plots
AF_landscape_plots <- plot_grid(
  AF_plot_Diam.Clutch + theme(legend.position = "none", 
                              axis.title.x = element_blank(),
                             axis.text.x = element_blank()) +
   # stat_contour(aes(z = average), breaks = breaks, linetype = "dotted", color = "black") + # mean(RF_Diam.Pref$absolute_fitness$average)
    geom_segment(data = gradients_Diam.Clutch, 
                 aes(x = sc.Diam, y = sc.log.Clutch, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))),
  AF_plot_Clutch.Pref + theme(legend.position = "none", 
                              axis.title.y = element_blank(),
                              axis.text.y = element_blank()) + xlab("Preference (SDs)") +
    stat_contour(aes(z = average), breaks = breaks, linetype = "dotted", color = "black") +
    geom_segment(data = gradients_Clutch.Pref, 
                 aes(x = sc.sqrt.Pref, y = sc.log.Clutch, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))), 
  AF_plot_Diam.Pref + theme(legend.position = "none") + ylab("Preference (SDs)") +
    #stat_contour(aes(z = average), breaks = breaks, linetype = "dotted", color = "black") +
    geom_segment(data = gradients_Diam.Pref, 
                 aes(x = sc.Diam, y = sc.sqrt.Pref, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))), 
  nrow=2, align="hv", labels = "AUTO")

AF_landscape_2d <- ggdraw(AF_landscape_plots) + draw_grob(AF_landscape_legend, x = 0.7, y=-0.2)
AF_landscape_2d
save_plot(filename = "MV_landscapes_2.pdf", plot = AF_landscape_2d, base_width=8, base_height = 6)
```

<!--
# Effect of Food-Web Treatment on Larval Survival

First we fit the full generalized linear mixed model to our data. This will enable us to evaluate whether there is any evidence that food-web treatment alters nonlinear selection on gall phenotypic variation. Note that we are not interested in directional selection yet, because we need to remove the nonlinear terms to evaluate its effects ([Lande and Arnold 1983](https://doi.org/10.1111/j.1558-5646.1983.tb00236.x)).

```{r Model 1, cache=TRUE, eval=FALSE, include=FALSE}
model_1 <- glmer(
  gall_survival ~ Foodweb + sc.Diam + sc.log.Clutch + sc.sqrt.Pref +
    Foodweb:(sc.Diam + sc.log.Clutch + sc.sqrt.Pref) +
    I(sc.Diam^2) + I(sc.log.Clutch^2) + I(sc.sqrt.Pref^2) +
    Foodweb:(I(sc.Diam^2) + I(sc.log.Clutch^2) + I(sc.sqrt.Pref^2)) +
    sc.Diam:sc.log.Clutch + sc.Diam:sc.sqrt.Pref + sc.log.Clutch:sc.sqrt.Pref +
    Foodweb:(sc.Diam:sc.log.Clutch + sc.Diam:sc.sqrt.Pref + sc.log.Clutch:sc.sqrt.Pref) +
    (1|Genotype/Plant_Position/Gall_Number),
  data = gall_selection.df,
  family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))
```

We use parametric bootstrapping (n=`r n_boots_analysis`) to evaluate the statistical significance of the effect of food-web treatment on nonlinear terms (i.e. Foodweb:trait~i~^2^ or Foodweb:trait~i~:trait~j~). Food-web treatment does influence the quadratic effect of of Oviposition preference on larval survival. None of the other terms statistically differ from zero.

```{r Bootstrap Model 1, cache=TRUE, include=FALSE, eval=FALSE}
boot_model_1 <- bootMer(model_1, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_1 <- tidy(model_1, conf.int=T) %>% #tidy(boot_model_1, conf.int=T) %>% 
  # filter terms that only matter for this analysis
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref",
                     "FoodwebSimple:I(sc.Diam^2)","FoodwebSimple:I(sc.log.Clutch^2)","FoodwebSimple:I(sc.sqrt.Pref^2)",
                     "FoodwebSimple:sc.Diam:sc.log.Clutch","FoodwebSimple:sc.Diam:sc.sqrt.Pref","FoodwebSimple:sc.log.Clutch:sc.sqrt.Pref")) 
```

```{r Plot Effect of Food-web on Nonlinear Terms in Model 1, echo=FALSE, eval=FALSE, include=FALSE}
plot_model_1 <- tidy_model_1 %>%
  mutate(Food_web = ifelse(grepl("FoodwebSimple", .$term)==TRUE,"Simple","Complex")) %>%
  filter(Food_web == "Simple")

plot_model_1$term_plot <- factor(plot_model_1$term,
                                levels=c("FoodwebSimple:I(sc.Diam^2)","FoodwebSimple:I(sc.log.Clutch^2)","FoodwebSimple:I(sc.sqrt.Pref^2)","FoodwebSimple:sc.Diam:sc.log.Clutch","FoodwebSimple:sc.Diam:sc.sqrt.Pref","FoodwebSimple:sc.log.Clutch:sc.sqrt.Pref"),
                                ordered=TRUE)

plot_model_1 %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), shape=21, size=3, fill="grey") + # y=statistic for bootstrap
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(plot_model_1$term_plot))) +
  ylab("Coefficient (logit scale)")
```

Therefore, we are going to drop them from the model. Note that we are still fitting the full curvature of the fitness landscape, but that we aren't allowing some aspects of it to vary between food-web treatments (since there was no evidence for it)

```{r Model 2, cache=TRUE, eval=FALSE, include=FALSE}

# Remove non-signficiant effect of food-web treatment on quadratic and trait_i:trait_j interactions

model_2 <- update(model_1, 
                  .~. -Foodweb:I(sc.Diam^2) -Foodweb:I(sc.log.Clutch^2) -Foodweb:I(sc.sqrt.Pref^2)
                  -Foodweb:sc.Diam:sc.log.Clutch -Foodweb:sc.Diam:sc.sqrt.Pref -Foodweb:sc.log.Clutch:sc.sqrt.Pref)
```

```{r Bootstrap Model 2, cache=TRUE, include=FALSE, eval=FALSE, include=FALSE}

boot_model_2 <- bootMer(model_2, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_2 <- tidy(model_2, conf.int=T) %>% # tidy(boot_model_2, conf.int=T) %>% 
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref",
                     "FoodwebSimple:I(sc.sqrt.Pref^2)")) 
```

This represents our nonlinear regression coefficients.

```{r Plot Nonlinear Terms, echo=FALSE, eval=FALSE, include=FALSE}
tidy_model_2$term_plot <- factor(tidy_model_2$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
                            ordered = TRUE)

tidy_model_2 %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), fill="grey", shape=21, size=3) + # y=statistic
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(tidy_model_2$term_plot))) +
  ylab("Coefficient (logit scale)")
```

Note that the non-overlapping 95% CI for *FoodwebSimple:I(sc.sqrt.Pref^2)* indicates that food-web treatment alters quadractic selection on Oviposition preference; however, this coefficient is not useful for actually calculating the selection gradient. To do this, we refit a model so rather than estimating the change in slope (currently), it will actually estimate the slope. See [Schielzeth 2010](https://doi.org/10.1111/j.2041-210X.2010.00012.x) for a detailed explanation of this approach.
```{r Adjust Model 2 for Slopes, eval=FALSE, include=FALSE}
model_2_adj <- update(model_2, .~. -I(sc.sqrt.Pref^2))
```

```{r Boot Adjusted Model 2 for accurate slope estimates, cache=TRUE, eval=FALSE, include=FALSE}
boot_model_2_adj <- bootMer(model_2_adj, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_2_adj <- tidy(model_2_adj, conf.int=T) %>% 
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebComplex:I(sc.sqrt.Pref^2)","FoodwebSimple:I(sc.sqrt.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"))
```

Importantly we see now that although quadratic selection varies between treatments, it is not different from zero in the simple food web.

```{r Plot Nonlinear Terms on appropriate scale, echo=FALSE, eval=FALSE, include=FALSE}
tidy_model_2_adj$term_plot <- factor(tidy_model_2_adj$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebComplex:I(sc.sqrt.Pref^2)","FoodwebSimple:I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
                            ordered = TRUE)

tidy_model_2_adj %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), fill="grey", shape=21, size=3) +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(tidy_model_2_adj$term_plot))) +
  ylab("Coefficient (logit scale)")
```

Now we fit a reduced model to assess the effects of whether food-web treatments alters components of directional selection

```{r Fit Model 3, cache=TRUE, eval=FALSE, include=FALSE}
model_3 <- update(model_2, 
                  .~. -I(sc.Diam^2) -I(sc.log.Clutch^2) -I(sc.sqrt.Pref^2) #-Foodweb:I(sc.sqrt.Pref^2) 
                  -sc.Diam:sc.log.Clutch -sc.Diam:sc.sqrt.Pref -sc.log.Clutch:sc.sqrt.Pref)
```

We then use parametric bootstrapping to test whether food-web treatment alters the linear relationship between each trait and larval survival.

```{r Bootstrap Model 3, cache=TRUE, include=FALSE, eval=FALSE}
boot_model_3 <- bootMer(model_3, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_3 <- tidy(model_3, conf.int=T) # boot_model_3
```

```{r Plot Model 3, echo=FALSE, eval=FALSE, include=FALSE}
plot_model_3 <- tidy_model_3 %>%
  mutate(Food_web = ifelse(grepl("FoodwebSimple:", .$term)==TRUE,"Simple","Complex")) %>%
  filter(Food_web == "Simple")

plot_model_3$term_ord <- factor(plot_model_3$term, 
                                levels=c("FoodwebSimple:sc.Diam","FoodwebSimple:sc.log.Clutch","FoodwebSimple:sc.sqrt.Pref"),
                                ordered = TRUE)
plot_model_3 %>%
  ggplot(., aes(x=term_ord)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), size=3, shape=21, fill="grey") + # y=statistic
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(limits=rev(levels(plot_model_3$term_ord))) +
  ylab("Coefficient (logit scale)")
```

There is clear evidence that food-web treatment alters the linear effect of clutch size on larval survival, but it has no effect on chamber diameter or Oviposition preference. Therefore, we fit a reduced model that removes these non-significant effects:

```{r Fit Model 4, cache=TRUE, eval=FALSE, include=FALSE}
model_4 <- update(model_3, .~. -Foodweb:sc.Diam -Foodweb:sc.sqrt.Pref)
```

This reduced model returns the same qualitative result as before. Specifically, that the linear effect of clutch size depends on food-web treatment. Note also that larval survival is significantly higher in the simple food web compared to the complex food web (dotted line). These are the regression coefficients we will use when calculating directional selection gradients ($\beta$).

```{r Bootstrap Model 4, cache=TRUE, include=FALSE, eval=FALSE, include=FALSE}
boot_model_4 <- bootMer(model_4, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_4 <- tidy(model_4, conf.int=T) # boot_model_4
```

```{r Plot Model 4, echo=FALSE, eval=FALSE, include=FALSE}
tidy_model_4$term_plot <- factor(tidy_model_4$term, 
                                levels = c("(Intercept)","FoodwebSimple","sc.Diam","sc.log.Clutch","FoodwebSimple:sc.log.Clutch","sc.sqrt.Pref"),
                                ordered = TRUE)

plot_model_4 <- tidy_model_4 %>%
  filter(term != "(Intercept)") %>%
  droplevels()

plot_model_4 %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), size=3, shape=21, fill="grey") + # y=statistic
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(plot_model_4$term_plot))) +
  ylab("Coefficient (logit scale)")
```

As we did with model 2, we refit the model by removing the slope term so that we can fit different slopes for each interaction term. 
```{r Fit Model 4 adjusted for estimating clutch slopes, cache=TRUE, eval=FALSE, include=FALSE}
model_4_adj <- update(model_4, .~. -sc.log.Clutch)
```

```{r Bootstrap adjusted model 4, cache=TRUE, include=FALSE, eval=FALSE}
boot_model_4_adj <- bootMer(model_4_adj, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_4_adj <- tidy(model_4_adj, conf.int=T) # boot_model_4_adj
```

```{r Plot adjusted model 4, echo=FALSE, eval=FALSE, include=FALSE}
tidy_model_4_adj$term_plot <- factor(tidy_model_4_adj$term, 
                                levels = c("(Intercept)","FoodwebSimple","sc.Diam","FoodwebComplex:sc.log.Clutch","FoodwebSimple:sc.log.Clutch","sc.sqrt.Pref"),
                                ordered = TRUE)

plot_model_4_adj <- tidy_model_4_adj %>%
  filter(term != "(Intercept)") %>%
  droplevels()

plot_model_4_adj %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate), size=3, shape=21, fill="grey") + # y=statistic
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(plot_model_4_adj$term_plot))) +
  ylab("Coefficient (logit scale)")
```

```{r Average larval survival, cache=TRUE, eval=FALSE, include=FALSE}
model_4_survival_ests <- update(model_4_adj, .~. -1)

boot_model_4_survival_ests <- bootMer(model_4_survival_ests, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_model_4_survival_ests <- tidy(model_4_survival_ests, conf.int=T) # boot_model_4_survival_ests

round(inverse_logit(filter(tidy_model_4_survival_ests, term=="FoodwebComplex")$estimate),2) # statistic
round(inverse_logit(filter(tidy_model_4_survival_ests, term=="FoodwebSimple")$estimate),2) # statistic
```

So, for plotting the results, we are going to use the focused nonlinear model. This provides the best representation of the fitness landscape (although we can't use it to assess to quantify the effects of directional selection.

Note that while there is no evidence to let the slopes vary among treatments, there is a clear effect of food web treatment on fitness. This could influence estimates of selection by magnifying them in the complex vs. simple food web.

So to calculate the selection gradients, we used Janzen and Stern's method to transform the regression coefficients.

Basically, I need to just multiply all of the coefficients by the simple and complex fitness. Except, however, for I(sc.sqrt.Pref) and sc.log.Clutch. Here, we need the different coefficients.

# Selection Gradients

We used the method of [Janzen and Stern (1998)](https://doi.org/10.1111/j.1558-5646.1998.tb02237.x) to calculate selection gradients. Specifically, we first multiply each regression coefficient (and confidence interval) by the average of $W(z)[1-W(z)]$, where $W(z)$ corresponds to the predicted absolute fitness of an individual (in this case 1 if the larva survived, 0 if not) given its multivariate phenotype ($z$). We then divided this value by the mean absolute fitness ($\bar W$) so our results are in terms of relative fitness ($w$). Note that we doubled all quadratic selection gradients ($\gamma_{z_i,z_i}$) to put them on the same scale as the directional and correlational selection gradients (detailed explanation in [Stinchcombe et al. 2008](https://doi.org/10.1111/j.1558-5646.2008.00449.x)).

```{r Organize and Transform Coefficients, include=FALSE, eval=FALSE}
tidy_coefs <- bind_rows(
  select(tidy_model_2, term, estimate, conf.low, conf.high), # estimate=statistic; tidy_model_2_adj no longer needed, since quadratic and correlational selection didn't differ among treatments
  select(tidy_model_4_adj, term, estimate, conf.low, conf.high)) %>% # estimate=statistic
  filter(term != "(Intercept)", term != "FoodwebSimple") # remove main effects

# Tidy common coefficients
tidy_common_coefs <- tidy_coefs %>%
  filter(#term!="FoodwebComplex:I(sc.sqrt.Pref^2)", term!="FoodwebSimple:I(sc.sqrt.Pref^2)",
         term!="FoodwebComplex:sc.log.Clutch",term!="FoodwebSimple:sc.log.Clutch")

# Tidy unique coefficients to each food-web
#complex_quad_pref <- filter(tidy_coefs, term %in% "FoodwebComplex:I(sc.sqrt.Pref^2)")
#complex_quad_pref$term <- "I(sc.sqrt.Pref^2)"
#simple_quad_pref <- filter(tidy_coefs, term %in% "FoodwebSimple:I(sc.sqrt.Pref^2)")
#simple_quad_pref$term <- "I(sc.sqrt.Pref^2)"

complex_clutch <- filter(tidy_coefs, term %in% "FoodwebComplex:sc.log.Clutch")
complex_clutch$term <- "sc.log.Clutch"
simple_clutch <- filter(tidy_coefs, term %in% "FoodwebSimple:sc.log.Clutch")
simple_clutch$term <- "sc.log.Clutch"
```

```{r Calculate Selection Gradients, include=FALSE, eval=FALSE}

# Estimate mean fitness and mean "brackets" for each food-web treatment (see Janzen and Stern 1998 equation 4 for details about "brackets")
complex_predict <- predict(model_2, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
complex_mean_brackets <- mean(complex_predict * (1 - complex_predict))
complex_mean_fitness <- mean(complex_predict)

simple_predict <- predict(model_2, newdata=filter(gall_selection.df, Foodweb=="Simple"), type="response")
simple_mean_brackets <- mean(simple_predict * (1 - simple_predict))
simple_mean_fitness <- mean(simple_predict)

# Double-check that using a different model (e.g. model_4) doesn't influence estimes of mean "brackets" and mean fitness. Below, I test this using the "Complex" food-web treatment as an example, but it doesn't matter 
alt_predict <- predict(model_4, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
alt_mean_brackets <- mean(alt_predict * (1 - alt_predict))
alt_mean_fitness <- mean(alt_predict)
test_mean_brackets <- complex_mean_brackets
test_mean_fitness <- complex_mean_fitness
test_mean_brackets - alt_mean_brackets # virtually the same
test_mean_fitness - alt_mean_fitness # virtually the same

tidy_simple_grads <- bind_rows(tidy_common_coefs, simple_clutch) %>% # simple_quad_pref, 
  mutate(Multiplier = ifelse(grepl("\\^2", .$term)==TRUE, 2, 1)) %>% # double all quadratic gradients so they are on the same scale as other directional and correlational selection gradients (detailed explanation in Stinchcombe et al. 2008)
  transmute(term=term,
            gradient = simple_mean_brackets * estimate / simple_mean_fitness * Multiplier,
            `2.5%` = simple_mean_brackets * conf.low / simple_mean_fitness * Multiplier,
            `97.5%` = simple_mean_brackets * conf.high / simple_mean_fitness * Multiplier)

tidy_complex_grads <- bind_rows(tidy_common_coefs, complex_clutch) %>% # complex_quad_pref, 
    mutate(Multiplier = ifelse(grepl("\\^2", .$term)==TRUE, 2, 1)) %>% # double all quadratic gradients so they are on the same scale as other directional and correlational selection gradients (detailed explanation in Stinchcombe et al. 2008)
  transmute(term=term,
            gradient = complex_mean_brackets * estimate / complex_mean_fitness * Multiplier,
            `2.5%` = complex_mean_brackets * conf.low / complex_mean_fitness * Multiplier,
            `97.5%` = complex_mean_brackets * conf.high / complex_mean_fitness * Multiplier)

selection_gradients <- bind_rows(
  mutate(tidy_simple_grads, `Food web`="Simple"),
  mutate(tidy_complex_grads, `Food web`="Complex")) #%>%
  #mutate(Type = ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic", ifelse(grepl(":", .$term)==TRUE, "Correlational", "Directional")))
```

```{r Table of Selection Gradients, echo=FALSE, eval=FALSE, include=FALSE}
selection_gradients$term_ord <- factor(selection_gradients$term, 
                                   levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref","I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
                                   labels=c("Diam","Clutch","Pref","Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"),
                                   ordered = TRUE)

selection_gradients_table <- selection_gradients %>%
  transmute(Term=term_ord, `Food web`, Gradient=round(gradient,2), `2.5%`=round(`2.5%`,2), `97.5%`=round(`97.5%`,2)) %>% 
  arrange(Term)

knitr::kable(selection_gradients_table, caption = "Table of Selection Gradients")
```

```{r Plot of Selection Gradients, echo=FALSE, eval=FALSE, include=FALSE}
selection_gradients %>%
  ggplot(., aes(x=term_ord)) +
  geom_linerange(aes(ymin=`2.5%`, ymax=`97.5%`, color=`Food web`), position=position_dodge(width=0.3)) +
  geom_point(aes(y=gradient, fill=`Food web`), color="black", shape=21, size=3, position=position_dodge(width=0.3)) +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(selection_gradients$term_ord))) +
  ylab("Selection Gradient") +
  scale_fill_manual(values=treatment_colors) +
  scale_color_manual(values=treatment_colors)
```

Contributions of larval parasitoids to selection gradients

To identify the mechanisms by which food-web treatment altered selection gradients, we refit our models so that the complex food-web treatment only contained cases of parasitisms from larval parasitoids. Therefore, these analyses identify the unique contribution of each parasitoid guild to the selection gradients (egg parasitoid from simple food web; larval parasitoid from complex food web). Note that cases of larval parasitism may have been initially parasitized by an egg parasitoid.

We restrict our model analyses to those where we observed different effects of food-web treatment on larval survival from the prior analyses (i.e. model 2 and model 4).

```{r Egg vs. Larval Parasitoids Model, cache=TRUE, eval=FALSE, include=FALSE}

# excludes cases of egg-parasitism from Complex food web
egg.vs.larval_df <- filter(gall_selection.df, Foodweb == "Simple" | Foodweb == "Complex" & platy < 1) %>% 
  mutate(Foodweb = ifelse(Foodweb == "Simple", "Egg", "Larval"))

egg.larval_model_2 <- update(model_2, data=egg.vs.larval_df)
```

```{r Bootstrap Egg vs. Larval Parasitoid Model 2, cache=TRUE, include=FALSE, eval=FALSE}

boot_egg.larval_model_2 <- bootMer(egg.larval_model_2, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_egg.larval_model_2 <- tidy(egg.larval_model_2, conf.int=T) %>% # boot_egg.larval_model_2
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"))#,
                     #"FoodwebLarval:I(sc.sqrt.Pref^2)")) 
```

This represents our nonlinear regression coefficients.

```{r Plot effect of Egg vs. Larval Parasitoid Model 2, echo=FALSE, eval=FALSE, include=FALSE}
tidy_egg.larval_model_2$term_plot <- factor(tidy_egg.larval_model_2$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
                            ordered = TRUE)

tidy_egg.larval_model_2 %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
   geom_point(aes(y=estimate), size=3, shape=21, fill="grey") +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(tidy_egg.larval_model_2$term_plot))) +
  ylab("Coefficient (logit scale)")
```
 To put all of the coefficients on the same scale, we refit it as we did in the food-web treatment
```{r Egg vs. Larval Parasitoids adjusted model 2, cache=TRUE, eval=FALSE, include=FALSE}
egg.larval_model_2_adj <- update(model_2_adj, data=egg.vs.larval_df)
```

```{r Bootstrap Egg vs. Larval Parasitoid adjusted Model 2, cache=TRUE, include=FALSE, eval=FALSE}

boot_egg.larval_model_2_adj <- bootMer(egg.larval_model_2_adj, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_egg.larval_model_2_adj <- tidy(egg.larval_model_2_adj, conf.int=T) %>% 
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebEgg:I(sc.sqrt.Pref^2)","FoodwebLarval:I(sc.sqrt.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) 
```

This represents our nonlinear regression coefficients.

```{r Plot effect of Egg vs. Larval Parasitoid adjusted Model 2, echo=FALSE, eval=FALSE, include=FALSE}
tidy_egg.larval_model_2_adj$term_plot <- factor(tidy_egg.larval_model_2_adj$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebEgg:I(sc.sqrt.Pref^2)","FoodwebLarval:I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
                            ordered = TRUE)

tidy_egg.larval_model_2_adj %>%
  ggplot(., aes(x=term_plot)) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
   geom_point(aes(y=estimate), size=3, shape=21, fill="grey") +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  scale_x_discrete(name="", limits = rev(levels(tidy_egg.larval_model_2_adj$term_plot))) +
  ylab("Coefficient (logit scale)")
```

Now we assess their contribution to the effect of food-web treatment on the linear terms.

```{r Fit Egg vs. Larval Parasitoid Model 4, cache=TRUE, eval=FALSE, include=FALSE}
egg.larval_model_4 <- update(model_4, data=egg.vs.larval_df)
```

This reduced model returns the same qualitative result as before. Specifically, that the linear effect of clutch size depends on food-web treatment. Note also that larval survival is significantly higher in the simple food web compared to the complex food web (dotted line). These are the regression coefficients we will use when calculating directional selection gradients ($\beta$).

```{r Bootstrap Egg vs. Larval Parasitoid Model 4, cache=TRUE, include=FALSE, eval=FALSE}
boot_egg.larval_model_4 <- bootMer(egg.larval_model_4, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

tidy_egg.larval_model_4 <- tidy(egg.larval_model_4, conf.int=T) # boot_egg.larval_model_4
```

Here, we observe clear evidence of conflicting selection pressures on clutch size. This results in net zero selection gradient acting on clutch size in the complex food web.
```{r Plot Egg vs. Larval Parasitoid Model 4, echo=FALSE, eval=FALSE, include=FALSE}
tidy_egg.larval_model_4$term_ord <- factor(tidy_egg.larval_model_4$term, 
                                levels = c("(Intercept)","FoodwebLarval","sc.Diam","sc.log.Clutch","FoodwebLarval:sc.log.Clutch","sc.sqrt.Pref"),
                                ordered = TRUE)

>>>>

---
title: "Reproduce Analyses Reported in "Phenotypic evolution is more constrained in simpler food webs""
author: "Matthew A. Barbour"
date: "`r Sys.Date()`"
output: github_document
# Trying github_document for easy viewing on github
#  prettydoc::html_pretty:
#    theme: architect
#    highlight: github
fontsize: 12pt
---

@@

## Load required libraries ----
#library(evolqg) # need to load before dplyr
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
#library(kableExtra)
library(stringr)
library(MVN)        # assessing multivariate normality assumption

@@

Reproduce Figure 3 in main manuscript.

@@
